Case based Arabic Morphological Analysis with Compositional Non-deterministic Automata

### 2010-06-25 trunk/tex/sarf.tex
finished related work

\usepackage{relsize}
\usepackage{fancyvrb}
%\usepackage{times}
\usepackage{latexsym}

@@

The hadith case study is interesting for Sarf since the controller
considers most of the analyses where we have a concentration
of proper names in the text, and ignores most of the analyses as long 
as they are valid where we do not have proper names. 
The controller is also interested in a small subset of words 
that relate persons to each other such as \RL{bin} or ``the sun of''
or words that mean narrate such as \RL{qaal} or ``said''. 
With Sarf, we successfully automated the analysis of 
three books of Islamic narrations~\cite{narrations1,narrations2, narrations3}
and extracted from each one of them a vector of a three level deep
complex structure.

@@

Sarf is similar to Beeseley~\shortcite{Beeseley:01} in that it uses
finiste-state automata (FSA). 
Beeseley reports that the number of the automata generated by a compiler
from Xerox rules is not controlled and also reports a difficulty to 
compose the FSAs into a single FSA framework. 
Sarf solves the problem with a non-deterministic composition of three
FSAs that are the prefix, stem, and suffix FSAs. 
With Sarf it is easy to control the FSAs with a user custom case 
controller, while with Beeseley the effort requires recompiling the
Xerox rules.

MAGEAD

SAMA: word based, treebank specific,

MADA+TOKAN: no controller, we do not need a tokenizer

AMIRA: machine learning...

ElixirFM:

@@

Sarf extends the lexicon of Buckwalter~\shotcite{Buckwalter:02} with 
proper names and location names extracted from different online 
sources~\footnote{\hyperref{http://alasmaa.net/}{http://alasmaa.net/ }, 
    \hyperref{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
as well as books 
\subsection{Recursive Affixes}
% smaller FSM
% retain POS info

>>>>

\usepackage{relsize}
\usepackage{fancyvrb}
%\usepackage{times}
\usepackage{latexsym}
\usepackage[colorlinks=false]{hyperref}

@@

The hadith case study is interesting for Sarf since the controller
considers most of the analyses where we have a concentration
of proper names in the text, and ignores most of the analyses as long 
as they are valid where we do not have proper names. 
The controller is also interested in a small subset of words 
that relate persons to each other such as \RL{bin} or ``the sun of''
or words that mean narrate such as \RL{qaal} or ``said''. 
With Sarf, we successfully automated the analysis of 
three books of Islamic narrations~\cite{IbnHanbal,AlTousi,AlKulayni}
and extracted from each one of them a vector of a three level deep
complex structure.

@@

Sarf is similar to Beesley~\shortcite{Beesley:01} in that it uses
finiste-state automata (FSA). 
Beesley reports that the number of the automata generated by a compiler
from Xerox rules is not controlled and also reports a difficulty to 
compose the FSAs into a single FSA framework. 
Sarf solves the problem with a non-deterministic composition of three
FSAs that are the prefix, stem, and suffix FSAs. 
With Sarf it is easy to control the FSAs with a user custom case 
controller, while with Beesley the effort requires recompiling the
Xerox rules.

Sarf is similar to The MADA+TOKAN toolkit in that it thinks of
the several valid morphological analyses as a richness that 
should be exploited at a higher abstract level rather than
an inaccuracy that should be corrected. 
Sarf differs in that it provides an interface for the 
case study to interfere in judging the analyses much earlier and
at a finer granularity level. 
Also in Sarf there is no need for a separate tokenizer such as
TOKAN as each alive automata carries with it the positions in 
text where a composition decision was made. 
The case-baded controller can build the trace when needed.

%MAGEAD 
    %Maamouri, Mohamed and \bgroup et al.\egroup },

We strongly feel that the refinemet of SAMA~\cite{Maamouri:10} 
is the closest to our approach. SAMA was refined to interact with
the ATB~\cite{Maamouri:04} project after the addition of a large 
new corpus. We find evidence in this decision that our hypothesis
about the utility of case-based analyzers. The refinements
that resulted in algorithmic changes in SAMA were done manually
and worked in integration with the ATB format only. 
In Sarf we propse a case-based controller approach that can modify 
the behaviour of the analyzer on the fly and that should interact
with the case study. In Sarf the analyzer remains general, and the 
case-based controller specializes the analyses.

The AMIRA~\cite{Diab:07} analyzer uses 
a language independent SVM based analysis. 
While the SVM analysis learns features from the text passed to the case
study, it does not make use of the type oh the query the 
case study targets in the text.

Like ElixirFM~\cite{Otakar:07} Sarf builds on the lexicon
of the Buckwalter analyzer. 
We also use deterministic parsing with tries or our analysis 
to run the prefix, suffix and stem FSAs. 
We think that their inferential-realizational approach 
that is highly compatible with the Arabic linguistic 
description~\cite{Badawi:04}
can make benefit of many features unique to the Arabic language.
Sarf leaves implementing that to the case-based controller
since in many cases the abstraction needed may not need the 
text to be grammar correct.
We can compare the case-based controller in Sarf to a high level
component in the ElixirFM domain sepcific language.

@@

Sarf extends the lexicon of Buckwalter~\shortcite{Buckwalter:02} with 
proper names and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
as well as biblical sources~\cite{}.

\subsection{Recursive Affixes}
% smaller FSM
% retain POS info
